\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-05 doc/paper.tex
doc/paper.tex: add section "Integrating custom C++ components"

In \fs the standard fixed-point iteration solver has been generalized
to handle towers of models (effective theories), which are matched at
intermediate scales.  An example of such a tower construction will be
given in \secref{sec:adapting-cpp-code}, where a right-handed neutrino
is integrated out at the see-saw scale, between the SUSY and the GUT
scale.

@@

\subsubsection{Changing boundary conditions}

@@

@@

@@

>>>>

In \fs the standard fixed-point iteration solver has been generalized
to handle towers of models (effective theories), which are matched at
intermediate scales.  An example of such a tower construction will be
given in \secref{sec:tower construction}, where right-handed neutrinos
are integrated out at the see-saw scale, between the SUSY and the GUT
scale.

@@

\subsubsection{Changing boundary conditions}
\label{sec:changing boundary conditions}

@@

\subsubsection{Stacking models in a tower of effective theories}
\label{sec:tower construction}

@@

\subsubsection{Integrating custom-built C++ components}

In \secref{sec:changing boundary conditions},
it was explained how one can let \fs generate
an alternative C++ boundary condition class
by authoring a model file.
Nonetheless, the way to employ this class at the C++ level
might still remain obscure to the reader
since \fs automatically took care of it.
Here, an example shall be exhibited
with the emphasis on the modular C++ code structure
that helps programming.
Concretely,
the auto-generated low-energy boundary condition on the MSSM
shall be modified so that
$\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}$
is determined from
$\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}$
by means of a two-loop matching.
This shall be accompanied by an improvement
of the $g_3$ $\beta$-function to the three-loop accuracy.

The first step is to alter
the model class which evaluates the $\beta$-functions.
Thanks to the \code{beta()} method being virtual,
one can override it conveniently by deriving a class from \code{MSSM}.
The declaration might look like:
% MSSMcbs_two_scale_model.hpp
\begin{numlstlisting}
#include "MSSM_two_scale_model.hpp"

template<>
class MSSMcbs<Two_scale> : public MSSM<Two_scale> {
public:
  explicit MSSMcbs(const MSSM_input_parameters& input_ = MSSM_input_parameters());
  virtual ~MSSMcbs();
  virtual Eigen::ArrayXd beta() const;
  MSSM_soft_parameters calc_beta() const;
};
\end{numlstlisting}
where the name \code{MSSMcbs} is an abbreviation of
the MSSM with custom-built $\beta$'s.
Note that the objects for \code{MSSM} are reused where possible:
\code{MSSM_input_parameters} in line 6 as well as
\code{MSSM_soft_parameters} in line 9.
This saves the programmer from excessive duplication of codes.
The member function definitions read:
\begin{numlstlisting}
Eigen::ArrayXd MSSMcbs<Two_scale>::beta() const
{
  return calc_beta().get();
}

MSSM_soft_parameters MSSMcbs<Two_scale>::calc_beta() const
{
  MSSM_soft_parameters betas(MSSM<Two_scale>::calc_beta());
  if (get_loops() <= 2) return betas;
  double bg33 = /* formula in terms of g1, g2, g3, Yu, Yd, Ye */;
  betas.set_g3(betas.get_g3() + Power(oneOver16PiSqr,3) * g3 * bg33);
  return betas;
}
\end{numlstlisting}
The full C++ expression of \code{bg33},
used in \code{MSSMcbs_two_scale_model.cpp},
has been adapted from the code by Jack and Jones \cite{3-loop MSSM betas}.
This already completes the amendment of the $g_3$ $\beta$-function.

The next step is to write a substitute for
the low-energy boundary condition class.
It must be declared as a descendent of \code{Constraint<Two_scale>}:
% MSSMcbs_two_scale_low_scale_constraint.hpp
\begin{numlstlisting}
#include "two_scale_constraint.hpp"

template<>
class MSSMcbs_low_scale_constraint<Two_scale> : public Constraint<Two_scale> {
public:
  MSSMcbs_low_scale_constraint(const MSSM_input_parameters&, const QedQcd&);
  virtual ~MSSMcbs_low_scale_constraint();
  void set_threshold_corrections(unsigned);
  ...
private:
  MSSMcbs<Two_scale>* model;
  QedQcd oneset;
  double new_g3;
  unsigned threshold_corrections;
  void calculate_DRbar_gauge_couplings();
  double calculate_alS5DRbar_over_alS5MSbar(double) const;
  double calculate_zeta_g_QCD_2(double) const;
  double calculate_zeta_g_SUSY_2(double) const;
  ...
In line 11,
the type of \code{model} has been adapted to the new model.
In fact, this class should work even if \code{model} remained
a pointer to \code{MSSM<Two_scale>} because of inheritance.
The main additions to \code{MSSM_low_scale_constraint}
in \code{models/MSSM/} are the member functions in lines 16--18,
which evaluate the two-loop matching coefficients
from Ref.~\cite{Harlander:2005wm}.
The following member function then performs
the two-step decoupling as reported in this reference:
% MSSMcbs_two_scale_low_scale_constraint.cpp
\begin{numlstlisting}
void MSSMcbs_low_scale_constraint<Two_scale>::calculate_DRbar_gauge_couplings()
  ...
  double alpha_s = oneset.displayAlpha(ALPHAS);
  double alS5DRbar_over_alS5MSbar = 1;
  double zeta_g_QCD_2 = 1;
  double zeta_g_SUSY_2 = 1;
  if (model->get_thresholds()) {
    alS5DRbar_over_als5MSbar = calculate_alS5DRbar_over_alS5MSbar(alpha_s);
    alpha_s *= alS5DRbar_over_alS5MSbar;  // alS5MSbar -> alS5DRbar
    zeta_g_QCD_2 = calculate_zeta_g_QCD_2(alpha_s);
    alpha_s /= zeta_g_QCD_2;		  // alS5DRbar -> alS6DRbar
    zeta_g_SUSY_2 = calculate_zeta_g_SUSY_2(alpha_s);
    alpha_s /= zeta_g_SUSY_2;		  // alS6DRbar -> alS6DRbarMSSM
    ...
  new_g3 = Sqrt(4*Pi * alpha_s);
  ...
one can integrate the new boundary condition class
\code{MSSMcbs_low_scale_constraint} together with
the new model \code{MSSMcbs}
into the spectrum generator in a straightforward manner.
They should supersede
\code{MSSM_low_scale_constraint} and \code{MSSM}, respectively.
The replacement should be carried out in
those objects that depend on these classes,
i.e.\ the initial guesser:
\begin{numlstlisting}
template<>
class MSSMcbs_initial_guesser<Two_scale> : public Initial_guesser<Two_scale> {
public:
  MSSMcbs_initial_guesser(MSSMcbs<Two_scale>*,
                          const MSSM_input_parameters&,
                          const QedQcd&,
                          const MSSMcbs_low_scale_constraint<Two_scale>&,
                          const MSSM_susy_scale_constraint<Two_scale>&,
                          const MSSM_high_scale_constraint<Two_scale>&);
  ...
private:
  MSSMcbs<Two_scale>* model;
  MSSM_input_parameters input_pars;
  QedQcd oneset;
  MSSMcbs_low_scale_constraint<Two_scale> low_constraint;
  MSSM_susy_scale_constraint<Two_scale> susy_constraint;
  MSSM_high_scale_constraint<Two_scale> high_constraint;
  ...
as well as the spectrum generator object:
% MSSMcbs_spectrum_generator.hpp
\begin{numlstlisting}
template <class T>
void MSSMcbs_spectrum_generator<T>::run(const QedQcd& oneset,
					const MSSM_input_parameters& input)
  ...
  MSSMcbs_initial_guesser<T> initial_guesser
    (&model, input, oneset,
     low_scale_constraint, susy_scale_constraint, high_scale_constraint);
  ...
The procedure described above is essentially all that is needed
to employ new spectrum generator components,
which may be composed from scratch or through
a linkage to external routines.
In doing so,
notice that there was an evident limit on the scope of
modules which one had to deal with.
For instance, it was clear from the outset
that one does not have to go through the code of the central
fixed-point iteration engine, \code{RGFlow}.
This manifests the power of the clean separation among objects
each with its well-defined distinct role.
One can find a working realization of this example
in \code{examples/customized-betas/}.

@@

\bibitem{3-loop MSSM betas}
  I.~Jack and D.~R.~T.~Jones, \ptitle{
    3-loop MSSM beta functions,}
  \url{http://www.liv.ac.uk/~dij/betas}.

\bibitem{Harlander:2005wm}
  R.~Harlander, L.~Mihaila and M.~Steinhauser,
  %``Two-loop matching coefficients for the strong coupling in the MSSM,''
  Phys.\ Rev.\ D {\bf 72} (2005) 095009
  [hep-ph/0509048].
  %%CITATION = HEP-PH/0509048;%%
  %27 citations counted in INSPIRE as of 05 Jun 2014
